### 2013-07-30 appa.tex
< 2 hours

\section{Computation for Map Counting}
\label{sec:count_map_computation}
\countmaps*
The main idea of the proof was given earlier on page~\pageref{thm:count_maps}.  Here are the details.

\begin{proof}[Proof of Proposition \ref{thm:count_maps}]
	First we resolve the case where $\ul r = \AA_n$, so that $\ell = n$ and $2^{\ell-n} = 1$.  We proceed by induction on $n \ge 1$ and consider eight cases.  (The base case is trivial.) The strategy for each is that, for the binary strings of length $n$ and top $T$, we consider those binary strings which end with $1$ and those which end with $0$.  Using \eqref{eq:top}, we may then write the new quantity in terms of the number of strings of length $n-1$ with a different top.

	When $n = 2k+1$, we have $\AA_n[n] = s$ and $n-1 = 2k$.  Let $\ul b'$ denote the first $n-1$ characters of $\ul b$.  Then,
	\newcommand{\aoeu}[5]{The top #1 ends in #2.  So, if $\ul b[n] = 1$ then we have top #3 for $\ul b'$, and otherwise $\ul b'$ has the same top #4.  In total, we get #5.}
	\begin{itemize}
		\ii \aoeu{$\AA_{2m+1}$}{$s$}{$\AA_{2m} = \AA_{2(m-1)+2}$}{$\AA_{2m+1}$}%
		{$\binom{n-2}{(m-1)+k} + \binom{n-2}{m+k} = \binom{n-1}{m+k}$}
		\ii \aoeu{$\AA_{2m}$}{$t$}{$\AA_{2m+1}$}{$\AA_{2m} = \AA_{2(m-1)+2}$}%
		{$\binom{n-2}{m+k} + \binom{n-2}{(m-1)+k} = \binom{n-1}{m+k}$}
		\ii \aoeu{$\AA_{2m-1}'$}{$t$}{$\AA_{2m}'$}{$\AA_{2m-1}' = \AA_{2(m-1)+1}'$}%
		{$\binom{n-2}{m+k} + \binom{n-2}{(m-1)+k} = \binom{n-1}{m+k}$}
		\ii \aoeu{$\AA_{2m}'$}{$s$}{$\AA_{2m-1}' = \AA_{2(m-1)+1}'$}{$\AA_{2m}'$}%
		{$\binom{n-2}{(m-1)+k} + \binom{n-2}{m+k} = \binom{n-1}{m+k}$}
	\end{itemize}
	The case where $n=2k$ is similar, but $n-1=2(k-1)+1$.  Now the last bit changes the top by $t$.
	\begin{itemize}
		\ii \aoeu{$\AA_{2m+1}$}{$s$}{$\AA_{2m+2} = \AA_{2(m+1)}$}{$\AA_{2m+1}$}%
		{$\binom{n-2}{(m+1)+(k-1)} + \binom{n-2}{m+(k-1)} = \binom{n-1}{m+k}$}
		\ii \aoeu{$\AA_{2m+2}$}{$t$}{$\AA_{2m+1}$}{$\AA_{2m+2} = \AA_{2(m+1)}$}%
		{$\binom{n-2}{m+(k-1)} + \binom{n-2}{(m+1)+(k-1)} = \binom{n-1}{m+k}$}
		\ii \aoeu{$\AA_{2m}'$}{$s$}{$\AA_{2m+1}' = \AA_{2(m+1)-1}'$}{$\AA_{2m}'$}%
		{$\binom{n-2}{(m+1)+(k-1)} + \binom{n-2}{m+(k-1)} = \binom{n-1}{m+k}$}
		\ii \aoeu{$\AA_{2m+1}'$}{$t$}{$\AA_{2m}'$}{$\AA_{2m+1}'=\AA_{2(m+1)-1}'$}%
		{$\binom{n-2}{m+(k-1)} + \binom{n-2}{(m+1)+(k-1)} = \binom{n-1}{m+k}$}
	\end{itemize}
	This completes the inductive step.  Hence, the proposition is true for $\ell = n$.

@@

\pasture*
\begin{proof}[Proof of Lemma \ref{thm:pasture}]
	Again, let $T$ denote the top, with length $N$.  In the first case, suppose $T[N] \neq \AA_n[n]$; that is, the last characters of $T$ and $\AA_n$ differ.  Then it must be the case that the rightmost anchor is not part of a fence.
	Now suppose the map arose from $\MM_{\AA_n}(\ul a, \ul b)$.  Let $\ul a'$ be the string formed by appending $N$ $1$'s to $\ul a$, and define $\ul b'$ similarly.  Now consider the map \[ \MM_{\AA_{n+N}} \left( \ul a', \ul b' \right). \]  See Figure~\ref{fig:pf_pasture} for an example.

@@

	The case where $T[N] = \AA_n[n]$ is similar.  We simply append a zero followed by $N$ $1$'s and repeat the same procedure.  The only difference is that the innermost barbell was generated by the $0$ we appended, and thus was not part of the original diagram.
\end{proof}

\section{Proof of the Alternating Recursion}
\label{sec:recurse_proof}
\recurse*
\begin{proof}[Proof of Proposition \ref{thm:recurse}]
	We will only resolve the case where $n$ is odd; that is, $\AA_n$ terminates in $s$.  The other case is identical.
	\begin{enumerate}[(i)]
		\ii Suppose $\ul a[n] = \ul b[n] = 0$.  Note that the top of $M_1$ is also $T$.
		\begin{itemize}
			\ii If $T = \varnothing$, then $M_1$ also had top $\varnothing$.  Therefore, the addition of the $0$'s at the end simply generates an extra blue barbell.  So the resulting map is simply a multiple by $\alpha_s$; i.e. $M = \alpha_sM_1$.
			\ii The case where $T$ ends in $t$ is analogous; we simply obtain a new barbell.  However, because $T \neq \varnothing$, the extra factor is $Q_T$ instead of simply $\alpha_t$.  Hence, $M = Q_TM_1$.
			\begin{figure}[ht]
				\centering
				\begin{asy}
					size(12cm);
					real h = 0.7;
					pen s = blue, t = red + dashed + 0.6;
					pen dot_s = blue, dot_t = red;
					int n = 4;

@@

					add(shift((7,0)) * three); add(shift((7,0)) * reflect((0,0),(1,0))*four);
					draw((7-1,0)--(7+5,0));
					MP("\to", (5,0), (0,0));
				\end{asy}
				\caption{$\MM(1000,0010)$ on the left, and $\MM(10000,00100)$ on right.}
				\label{fig:recurse_example_destr_zero}
			\end{figure}
			\ii Now suppose $T$ ends in $s$.  Then the addition of the $0$'s resulted in enclosing a caterpillar of $M_1$ in an attached bubble, as per Figure~\ref{fig:recurse_example_destr_zero}.  Breaking out of the additional blue loop nows adds a factor of $x$, but the red caterpillar is annihilated in the process; therefore, only the fluff remains, because there is nothing to push through the fences.  Hence, $M = -xF$.
		\end{itemize}
		\ii Now suppose that $\ul a[n] = \ul b[n] = 1$.
		\begin{itemize}
			\ii First, suppose $T$ ends in $s$.  Then the additional $1$'s added simply create a fence on the far right, because there is no ``up'' blue anchors to interact with.  Because this fence is already on the far right, it does not affect the product, hence $M = M_1$.
			\begin{figure}[ht]
				\centering
				\begin{asy}
					size(11cm);
					real h = 0.7;
					pen s = blue, t = red + dashed + 0.6;
					pen dot_s = blue, dot_t = red;
					int n = 4;

@@

					add(shift((7,0)) * three); add(shift((7,0)) * reflect((0,0),(1,0)) * four);
					draw((7-1,0)--(7+5,0));
					MP("\to", (5,0), (0,0));
				\end{asy}
				\caption{$\MM(0110,0110)$ on the left, and $\MM(01101,01101)$ on right.}
				\label{fig:recurse_example_destr_one}
			\end{figure}
			\ii Conversely, suppose that $T$ ends in $t$.   See Figure \ref{fig:recurse_example_destr_one} for an example.

			In that case, $M_1$ must have had the top $T$ with an $s$ appended.  Once again, we find that the red caterpillar from $M_1$ is enclosed in a bubble. This time, however, the bubble is not attached.  Therefore, while we still obtain the extra factor of $x$ and the caterpillar is still removed, we remain with a blue barbell on the far right pasture.  So we must push this blue barbell through $T$.  This adds a factor of $Q_T$, and so we obtain the desired form $M = -xF Q_T$.
			\ii The case where $T = \varnothing$ is analogous to the previous case; in a sense, we merely have a degenerate $Q_\varnothing = \alpha_s$.  Specifically, the blue barbell is already on the correct pasture, so we simply add a factor of $\alpha_s$.  Therefore, $M = -xF \alpha_s$.
		\end{itemize}
	\end{enumerate}
	This exhausts all the cases in the proposition, hence we're done.
\end{proof}

>>>>

\section{Computation for Map Counting}
\label{sec:count_map_computation}
\defredcountmaps*
\countmaps*
The main idea of the proof was given earlier on page~\pageref{thm:count_maps}.  Here are the details.

\begin{proof}[Proof of Proposition \ref{thm:count_maps}]
	First we resolve the case where $\ul r = \AA_n$, so that $\ell = n$ and $2^{\ell-n} = 1$.  We proceed by induction on $n \ge 1$ and consider eight cases.  (The base case is trivial.) The strategy for each is that, for the binary strings of length $n$ and top $\ul w$, we consider those binary strings which end with $1$ and those which end with $0$.  Using \eqref{eq:top}, we may then write the new quantity in terms of the number of strings of length $n-1$ with a different top.

	When $n = 2k+1$, we have $\AA_n[n] = s$ and $n-1 = 2k$.  Let $\ul b'$ denote the first $n-1$ characters of $\ul b$.  Then,
	\newcommand{\aoeu}[5]{The top #1 ends in #2.  So, if $\ul b[n] = 1$ then we have top #3 for $\ul b'$, and otherwise $\ul b'$ has the same top #4.  In total, we get #5.}
	\begin{itemize}
		\ii \aoeu{$\AA_{2m+1}$}{$s$}{$\AA_{2m} = \AA_{2(m-1)+2}$}{$\AA_{2m+1}$}%
		{$\binom{n-2}{(m-1)+k} + \binom{n-2}{m+k} = \binom{n-1}{m+k}$}
		\ii \aoeu{$\AA_{2m}$}{$t$}{$\AA_{2m+1}$}{$\AA_{2m} = \AA_{2(m-1)+2}$}%
		{$\binom{n-2}{m+k} + \binom{n-2}{(m-1)+k} = \binom{n-1}{m+k}$}
		\ii \aoeu{$\AA_{2m-1}'$}{$t$}{$\AA_{2m}'$}{$\AA_{2m-1}' = \AA_{2(m-1)+1}'$}%
		{$\binom{n-2}{m+k} + \binom{n-2}{(m-1)+k} = \binom{n-1}{m+k}$}
		\ii \aoeu{$\AA_{2m}'$}{$s$}{$\AA_{2m-1}' = \AA_{2(m-1)+1}'$}{$\AA_{2m}'$}%
		{$\binom{n-2}{(m-1)+k} + \binom{n-2}{m+k} = \binom{n-1}{m+k}$}
	\end{itemize}
	The case where $n=2k$ is similar, but we now have $n-1=2(k-1)+1$.  Now the last bit changes the top by $t$.
	\begin{itemize}
		\ii \aoeu{$\AA_{2m+1}$}{$s$}{$\AA_{2m+2} = \AA_{2(m+1)}$}{$\AA_{2m+1}$}%
		{$\binom{n-2}{(m+1)+(k-1)} + \binom{n-2}{m+(k-1)} = \binom{n-1}{m+k}$}
		\ii \aoeu{$\AA_{2m+2}$}{$t$}{$\AA_{2m+1}$}{$\AA_{2m+2} = \AA_{2(m+1)}$}%
		{$\binom{n-2}{m+(k-1)} + \binom{n-2}{(m+1)+(k-1)} = \binom{n-1}{m+k}$}
		\ii \aoeu{$\AA_{2m}'$}{$s$}{$\AA_{2m+1}' = \AA_{2(m+1)-1}'$}{$\AA_{2m}'$}%
		{$\binom{n-2}{(m+1)+(k-1)} + \binom{n-2}{m+(k-1)} = \binom{n-1}{m+k}$}
		\ii \aoeu{$\AA_{2m+1}'$}{$t$}{$\AA_{2m}'$}{$\AA_{2m+1}'=\AA_{2(m+1)-1}'$}%
		{$\binom{n-2}{m+(k-1)} + \binom{n-2}{(m+1)+(k-1)} = \binom{n-1}{m+k}$}
	\end{itemize}
	This completes the inductive step.  Hence, the proposition is true for $\ell = n$.

@@

\pasture*
\begin{proof}[Proof of Lemma \ref{thm:pasture}]
	Again, let $\ul w$ denote the top, with length $N$.  In the first case, suppose $\ul w[N] \neq \AA_n[n]$; that is, the last characters of $\ul w$ and $\AA_n$ differ.  Then it must be the case that the rightmost anchor is not part of a fence.
	Now suppose the map arose from $\MM_{\AA_n}(\ul a, \ul b)$.  Let $\ul a'$ be the string formed by appending $N$ $1$'s to $\ul a$, and define $\ul b'$ similarly.  Now consider the map \[ \MM_{\AA_{n+N}} \left( \ul a', \ul b' \right). \]  See Figure~\ref{fig:pf_pasture} for an example.

@@

	The case where $\ul w[N] = \AA_n[n]$ is similar.  We simply append a zero followed by $N$ $1$'s and repeat the same procedure.  The only difference is that the innermost barbell was generated by the $0$ we appended, and thus was not part of the original diagram.
\end{proof}

\section{Proof of the Alternating Recursion}
\label{sec:recurse_proof}
\recurse*
\begin{proof}[Proof of Proposition \ref{thm:recurse}]
	We will only resolve the case where $n$ is odd; that is, $\AA_n$ terminates in $s$.  The other case is identical.
	\begin{enumerate}[(i)]
		\ii Suppose $\ul a[n] = \ul b[n] = 0$.  Note that the top of $M_1$ is also $\ul w$.
		\begin{itemize}
			\ii If $\ul w = \varnothing$, then $M_1$ also had top $\varnothing$.  Therefore, the addition of the $0$'s at the end simply generates an extra blue barbell.  So the resulting map is simply a multiple by $\alpha_s$; i.e. $M = \alpha_sM_1$.
			\ii The case where $\ul w$ ends in $t$ is analogous; we simply obtain a new barbell.  However, because $\ul w \neq \varnothing$, the extra factor is $Q_{\ul w}$ instead of simply $\alpha_t$.  Hence, $M = Q_{\ul w}M_1$.
			\begin{figure}[ht]
				\centering
				\begin{asy}
					size(12cm);
					real h = 0.7;
					pen s = blue, t = red + dashed + 0.6;
					pen dot_s = blue, dot_t = red;
					int n = 4;

@@

					add(shift((7,0)) * three); add(shift((7,0)) * reflect((0,0),(1,0))*four);
					draw((7-1,0)--(7+5,0));
					MP("\to", (5,0), (0,0));
				\end{asy}
				\caption{$\MM(1000,0010)$ on the left, and $\MM(10000,00100)$ on right.}
				\label{fig:recurse_example_destr_zero}
			\end{figure}
			\ii Now suppose $\ul w$ ends in $s$.  Then the addition of the $0$'s resulted in enclosing a caterpillar of $M_1$ in an attached bubble, as per Figure~\ref{fig:recurse_example_destr_zero}.  Breaking out of the additional blue loop nows adds a factor of $x$, but the red caterpillar is annihilated in the process; therefore, only the fluff remains, because there is nothing to push through the fences.  Hence, $M = -xF$.
		\end{itemize}
		\ii Now suppose that $\ul a[n] = \ul b[n] = 1$.
		\begin{itemize}
			\ii First, suppose $\ul w$ ends in $s$.  Then the additional $1$'s added simply create a fence on the far right, because there is no ``up'' blue anchors to interact with.  Because this fence is already on the far right, it does not affect the product, hence $M = M_1$.
			\begin{figure}[ht]
				\centering
				\begin{asy}
					size(11cm);
					real h = 0.7;
					pen s = blue, t = red + dashed + 0.6;
					pen dot_s = blue, dot_t = red;
					int n = 4;

@@

					add(shift((7,0)) * three); add(shift((7,0)) * reflect((0,0),(1,0)) * four);
					draw((7-1,0)--(7+5,0));
					MP("\to", (5,0), (0,0));
				\end{asy}
				\caption{$\MM(0110,0110)$ on the left, and $\MM(01101,01101)$ on right.}
				\label{fig:recurse_example_destr_one}
			\end{figure}
			\ii Conversely, suppose that $\ul w$ ends in $t$.   See Figure \ref{fig:recurse_example_destr_one} for an example.

			In that case, $M_1$ must have had the top $\ul w$ with an $s$ appended.  Once again, we find that the red caterpillar from $M_1$ is enclosed in a bubble. This time, however, the bubble is not attached.  Therefore, while we still obtain the extra factor of $x$ and the caterpillar is still removed, we remain with a blue barbell on the far right pasture.  So we must push this blue barbell through $\ul w$.  This adds a factor of $Q_T$, and so we obtain the desired form $M = -xF Q_{\ul w}$.
			\ii The case where $\ul w = \varnothing$ is analogous to the previous case; in a sense, we merely have a degenerate $Q_\varnothing = \alpha_s$.  Specifically, the blue barbell is already on the correct pasture, so we simply add a factor of $\alpha_s$.  Therefore, $M = -xF \alpha_s$.
		\end{itemize}
	\end{enumerate}
	This exhausts all the cases in the proposition, hence we're done.
\end{proof}
